Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-06-03 thesis/chapters/ch_chapter4.tex
edits to chapter4 - still need to fix figures

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patterns within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018} and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Duggan2018,Nagasawa2018,Ojima2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way exhibits smaller abundance spreads, averaging over the variety of complex hydrodynamics and feedback-driven turbulence that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns in low metallicity environments of the Milky Way halo \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. Indeed the increase in scatter in the stellar abundance patterns of low mass dwarf galaxies and UFDs has been attributed to inhomogeneous mixing and stochastic effects \citep[e.g.][]{Norris2010a,Lee2013,Simon2015,Mashonkina2017,Suda2017}. This complicates the interpretation of abundance patterns in these galaxies, particularly for constraining individual enrichment events.

@@

\section{Results}
\label{ch4:sec:results}
Perhaps the three most important parameters to quantify for each enrichment event are: 1) what fraction of released metals are immediately available for star formation, 2) what fraction of metals are carried out of the galaxy in outflows, and 3) how and over what timescale do metals cool from hot phases into star forming gas. We discuss each of these points in Section~\ref{ch4:sec:ISM CGM}. In Section~\ref{ch4:sec:spreads} we go further and address the relative homogeneity of the metals retained by the ISM for each enrichment source.

@@

Gas above the star formation threshold in our simulations is limited and short-lived due to the low gas surface densities and star formation rates in this galaxy and the efficiency of stellar feedback from newly formed stars. As a proxy, we examine the evolution of the CNM, out of which the star forming gas forms. In general, very few of these elements are available for immediate star formation in the CNM ($<< 1\%$, see Figure~\ref{ch4:fig:CGM_CNM}). However, we note that we may not have sufficient resolution to properly resolve the details of the initial mixing of individual enrichment events in the ISM. In addition, this value will be sensitive to whether or not a given event occurs in the vicinity of or inside an active, star forming region -- as is likely for core collapse SNe, and unlikely for other events. Thus, this is a quantity that does not depend solely on $E_{\rm ej}$ and is sensitive to whether or not self-enrichment of star forming regions is an important factor in galactic chemical evolution. We are also missing important physical processes, like dust production in AGB winds and core collapse SNe, that may affect these results for specific enrichment channels in ways that do not depend solely on $E_{\rm ej}$. Investigating the fraction of metals immediately available for star formation will require additional, high-resolution simulations of metal mixing in and around individual star forming regions.
% add references (maybe) to high res simulation of dust production in AGB winds and SNe ???

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers. Conducting this analysis requires measured r-process abundances for a large sample of stars in an individual dwarf galaxy. This is challenging for the lowest mass, most metal poor dwarf galaxies, which typically only have a few to 10 stars with measured abundances, but could be done for low mass dwarfs like Ursa Minor, and for more massive dwarfs like Carina, and Fornax \citep{Suda2017}.
% somerthing about needing cool onezone chemical models to try and model data with many free params might be nice here.

@@

 \begin{figure*}
   \centering
   \includegraphics[width=0.975\linewidth]{figures/ch4/II_Eej_CNM_avg}
   \caption{A summary of the results presented in Figure~\ref{ch4:fig:CGM_CNM} and Figure~\ref{ch4:fig:CGM_CNM} that demonstrates the significant variation in $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the mean-median separate at fixed injection energy. The points give the average values at each energy at two different times, while the error bars show the min/max values at each energy. The points at different times have been offset slightly along the horizontal axes for clarity. Since they are only two points, we show the results from \runone E52\_r0 and \runone E52\_r300 without averaging.}
   \label{ch4:fig:CGM_CNM_variance}
 \end{figure*}

>>>>

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patterns within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018} and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Duggan2018,Nagasawa2018,Ojima2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way exhibits smaller abundance spreads. The size of more massive galaxies reduces the relevance of stochasticity, averaging over the variety of complex hydrodynamics and feedback-driven turbulence that may be driving much of the significant vgariance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns in low metallicity environments of the Milky Way halo \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. Indeed the increase in scatter in the stellar abundance patterns of low mass dwarf galaxies and UFDs has been attributed to inhomogeneous mixing and stochastic effects \citep[e.g.][]{Norris2010a,Lee2013,Simon2015,Mashonkina2017,Suda2017}. This complicates the interpretation of abundance patterns in these galaxies, particularly for constraining individual enrichment events.

@@

\section{Results}
\label{ch4:sec:results}
Perhaps the three most important parameters to quantify for each enrichment event are: 1) what fraction of released metals are immediately available for star formation, 2) how does this fraction evolve over time as metals cool from hot phases into star forming gas, and 3) what fraction of metals are carried out of the galaxy in outflows. We discuss each of these points in Section~\ref{ch4:sec:ISM CGM}, focusing on the average behavior of multiple enrichment events at fixed ejection energy. In Section~\ref{ch4:sec:spreads} we go further and address the relative homogeneity (over time) of the metals retained by the ISM for each enrichment source. Finally, in Section~\ref{ch4:sec:individual event variation} we discuss how these parameters vary with individual events, and, in Sections~\ref{ch4:sec:radial position} through ~\ref{ch4:sec:SFR}, how they vary with radial position in the galaxy, ISM properties in the event region, and global galaxy SFR.

@@

Gas above the star formation threshold in our simulations is limited and short-lived due to the low gas surface densities and star formation rates in this galaxy and the efficiency of stellar feedback from newly formed stars. As a proxy, we examine the evolution of the CNM, out of which the star forming gas forms. In general, very few of these elements are available for immediate star formation in the CNM ($<< 1\%$, see Figure~\ref{ch4:fig:CGM_CNM}). However, we note that we may not have sufficient resolution to properly resolve the details of the initial mixing of individual enrichment events in the ISM. In addition, this value will be sensitive to whether or not a given event occurs in the vicinity of or inside an active, star forming region -- as is likely for core collapse SNe, and unlikely for other events. Thus, this is a quantity that does not depend solely on $E_{\rm ej}$ and is sensitive to whether or not self-enrichment of star forming regions is an important factor in galactic chemical evolution. We are also missing important physical processes, like dust production in AGB winds and core collapse SNe, that may affect these results for specific enrichment channels in ways that do not depend solely on $E_{\rm ej}$. Investigating the fraction of metals immediately available for star formation will require additional, high-resolution simulations of metal mixing in and around individual star forming regions. This will be useful for understanding stellar enrichment patterns that may arise from star formation in regions whose metal abundance patters arise directly from unmixed ejecta of a single enrichment source. % cite
% add references (maybe) to high res simulation of dust production in AGB winds and SNe ???

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers. Conducting this analysis requires measured r-process abundances for a large sample of stars in an individual dwarf galaxy. This is challenging for the lowest mass, most metal poor dwarf galaxies, which typically only have a few to 10 stars with measured abundances, but could be done for low mass dwarfs like Ursa Minor, for more massive dwarfs like Carina, and Fornax \citep{Suda2017}, and for the low-metallicity stellar halo of the Milky Way \citep[e.g.][]{Hansen2018,Sakari2018}.
% somerthing about needing cool onezone chemical models to try and model data with many free params might be nice here.

@@

 \begin{figure*}
   \centering
   \includegraphics[width=0.975\linewidth]{figures/ch4/II_Eej_CNM_avg}
   \caption{A summary of the results presented in Figure~\ref{ch4:fig:CGM_CNM} and Figure~\ref{ch4:fig:mean-median} that demonstrates the significant variation in $f_{\rm CGM}$, $f_{\rm CNM} / f_{\rm Disk}$, and the mean-median separate at fixed injection energy. The points give the average values at each energy at two different times, while the error bars show the min/max values at each energy. The points at different times have been offset slightly along the horizontal axes for clarity. Since they are only two points, we show the results from \runone E52\_r0 and \runone E52\_r300 without averaging.}
   \label{ch4:fig:CGM_CNM_variance}
 \end{figure*}
